---
title: "Lol"
author: "Ian Dinnie"
date: "1/14/2021"
output: word_document
---

```{r setup, include=T}
suppressMessages(library(tidyverse))

```

```{r}
# Things to add
# Splits
# I need to perfect iterating through entire deck: right now it is set to stop when there are 30 cards remaining, when the deck gets too small to go through it a whole iteration it throws up an error because it runs out of cards to sample
# More players

```

```{r}
#construct a deck
deck <- sort(rep(c(1,2,3,4,5,6,7,8,9,10),4)) # create a vector
rep10 <- rep(deck[37:40],3) # create a bunch of 10s
deck <- sort(rep(c(deck,rep10),8)) # combine into one deck of 416 cards

# write a function that updates the deck
update_deck <- function(x, y){
  pulled <- rep(y, sum(x == y) - 1)
  newdeck <- x[!x %in% y]
  newdeck <- sort(c(pulled, newdeck))
}

# write a function that checks ace to determine if it should be a 1 or 11
checkace <- function(a,b){
  if(a == 1){
    if(11 + b > 21) {
      a <- 1
    } else {
      a <- 11
    }
  }
  if(b == 1){
    if (11 + a > 21){
      b <- 1
    } else {
      b <- 11
    }
  } 
  c(a,b)
}

# make another one just for hits
checkacehit <- function(z, i){
  if(z == 1){
    if(i + 11 > 21) {
      z <- 1
    } else {
      z <- 11
    }
  }
  return(z)
}

# a function for splitting hands, not done yet
splitjack <- function(s){
  split1 <- c(dealp[1], sample(newdeck,1,replace = F))
  newdeck <- update_deck(newdeck, split1[2])
  deal <- checkace(split1[1], split1[2])
  add <- split1[1] + split1[2]
  while(add <= 16){ # if the sum is less than 16, hit, until at or above 16
    hit <- sample(newdeck, 1, replace = F)
    newdeck <- update_deck(newdeck, hit)   
    add = add + hit # sum add and hit, this process continues until it exceeds 16
  }
}

#the beginning; this algorithm draws two cards from the deck and proceeds to hit until at or above 16; this is the shell for how the whole system will work; need to add player(s) to play against that use same rules + the bust card rule (don't hit if dealer is showing 4,5,6)
blackjack <- function(x){
  deal <- sample(x, 2, replace = F) # deal 2 cards from the deck
  newdeck <- x %>% update_deck(deal)
  deal <- checkace(deal[1], deal[2])
  add <- deal[1] + deal[2]
  while(add <= 16){ # if the sum is less than 16, hit, until at or above 16
    hit <- sample(newdeck, 1, replace = F)
    newdeck <- update_deck(newdeck, hit)   
    add = add + hit # sum add and hit, this process continues until it exceeds 16
  } 
  return(add) # return the final sum
}
# it is easy to add another if qualifier to return the word 'bust' if it exceeds 21, but I want to use this for simulations/statistical analysis, so better to not add strings with characters
#test <- blackjack(deck) # test it :D
#runs <- rerun(100,blackjack(deck))
#runs <- as_vector(runs)
#mean(runs)
```

```{r}
# Add players

#construct a deck
deck <- sort(rep(c(1,2,3,4,5,6,7,8,9,10),4)) # create a vector
rep10 <- rep(deck[37:40],3) # create a bunch of 10s
deck <- sort(rep(c(deck,rep10),8))

twoplayerblackjack <- function(x){
  deald <- sample(x, 1, replace = F) # deal first card to dealer
  newdeck <- x %>% update_deck(deald) # update deck
  dealp <- sample(newdeck, 1, replace = F) # deal first card to player
  newdeck <- update_deck(newdeck, dealp) # update deck
  deald <- c(deald, sample(newdeck, 1, replace = F)) # deal second card to dealer
  newdeck <- update_deck(newdeck, deald[2]) # update deck
  deald <- checkace(deald[1], deald[2]) # check aces of dealer
  dealp <- c(dealp, sample(newdeck, 1, replace = F)) # deal second card to player
  newdeck <- update_deck(newdeck, dealp[2]) # update deck
  dealp <- checkace(dealp[1], dealp[2]) # check aces of player
 # we now have two cards dealt to the player and dealer (in the proper sequence), and an updated deck, yay
  
 # it is now the player's turn to assess their cards and hit or stay (want to add splits at some point)
  if(deald[1] == 4 | deald[1] == 5 | deald[1] == 6){ # check for bust cards, stay if showing, interestingly, this strat raised house odds 2 points when ran once, unclear if meaningful 
    addp <- dealp[1] + dealp[2]
  } else{
  addp <- dealp[1] + dealp[2]
   while(addp <= 16){ # if the sum is less than 16, hit, until at or above 16
    hit <- sample(newdeck, 1, replace = F)
    newdeck <- update_deck(newdeck, hit)
    hit <- checkacehit(hit, addp)
    addp = addp + hit # sum add and hit, this process continues until it exceeds 16
   }
  }
  # it is now the dealer's turn
  addD <- deald[1] + deald[2]
  while(addD <= 16){ # if the sum is less than 16, hit, until at or above 16
    hit <- sample(newdeck, 1, replace = F)
    newdeck <- update_deck(newdeck, hit)
    hit <- checkacehit(hit, addD)
    addD = addD + hit # sum add and hit, this process continues until it exceeds 16
  }
  
  #print(addp)
  #print(addD)
  #print(length(newdeck))
  if (addp > addD & addp < 22 | addD > 21 & addp <= 21){
    #print("Player wins")
    Result <- "Player Win"
  } else {
    if(addp < addD & addD < 22| addp > 21 & addD <= 21){
      #print("House Wins")
      Result <- "House Win"
    } else {
    #print("Draw")
      if(addp == addD | addp > 21 & addD > 21 )
      Result <- "Draw"
  }
  }
}

  # It works! 
```

```{r}
# Run through entire deck
twoplayerblackjack_test <- function(x){
  hand <- 0
  decklength <- length(x)
  while(decklength > 30){
  hand <- hand + 1
  deald <- sample(deck, 1, replace = F) # deal first card to dealer
  deck <- deck %>% update_deck(deald) # update deck
  dealp <- sample(deck, 1, replace = F) # deal first card to player
  deck <- update_deck(deck, dealp) # update deck
  deald <- c(deald, sample(deck, 1, replace = F)) # deal second card to dealer
  deck <- update_deck(deck, deald[2]) # update deck
  deald <- checkace(deald[1], deald[2]) # check aces of dealer
  dealp <- c(dealp, sample(deck, 1, replace = F)) # deal second card to player
  deck <- update_deck(deck, dealp[2]) # update deck
  dealp <- checkace(dealp[1], dealp[2]) # check aces of player
 # we now have two cards dealt to the player and dealer (in the proper sequence), and an updated deck, yay
  
 # it is now the player's turn to assess their cards and hit or stay (want to add splits at some point)
  if(deald[1] == 4 | deald[1] == 5 | deald[1] == 6){ # check for bust cards, stay if showing
    addp <- dealp[1] + dealp[2]
  } else {
  addp <- dealp[1] + dealp[2]
   while(addp <= 16){ # if the sum is less than 16, hit, until at or above 16
    hit <- sample(deck, 1, replace = F)
    deck <- update_deck(deck, hit)
    hit <- checkacehit(hit, addp)
    addp = addp + hit # sum add and hit, this process continues until it exceeds 16
   }
  }
  # it is now the dealer's turn
  addD <- deald[1] + deald[2]
  while(addD <= 16){ # if the sum is less than 16, hit, until at or above 16
    hit <- sample(deck, 1, replace = F)
    deck <- update_deck(deck, hit)
    hit <- checkacehit(hit, addD)
    addD = addD + hit # sum add and hit, this process continues until it exceeds 16
  }
  
  #print(addp)
  #print(addD)
  #print(length(deck))
  if (addp > addD & addp < 22 | addD > 21 & addp <= 21){
    Result <- "Player Win"
  } else {
    if(addp < addD & addD < 22| addp > 21 & addD <= 21){
      Result <- "House Win"
    } else {
      if(addp == addD | addp > 21 & addD > 21 )
      Result <- "Draw"
  }
  }
    decklength <- length(deck)  
    if(hand == 1){
      Result_table <- tibble(Result)
    } else {
      new <- tibble(Result)
      Result_table <- rbind(Result_table, new)
    }
  }
  return(Result_table)
}
# A full-ish game of blackjack between a dealer and player!!!
```

```{r}
# simulate 1000 games
# Result <- twoplayerblackjack(deck)
sim <- rerun(1000,twoplayerblackjack_test(deck)) # this works but need to figure out how to use list output

allresults <- unlist(sim) # get a vector of all results across all hands 
player <- mean(allresults == "Player Win")
house <- mean(allresults == "House Win")
draw <- mean(allresults == "Draw")

ResultsofSim <- matrix(data = 100*c(player, house, draw), nrow = 1, ncol = 3, dimnames = list("%",c("Player Win %", "House Win %", "Draw %")), byrow = T) # create a neat matrix to display sim results
print(ResultsofSim)


#Averages_test <- test_sim %>% 
 # mutate(dummy = rep(1,10000)) %>%
  #group_by(value) %>% 
  #summarise(count = sum(dummy == 1)) %>% 
  #mutate(Share = count/sum(count))
```



























#### Archive #####
``{r}
# OLD # 
# trying to add another player

#construct a deck
deck <- sort(rep(c(1,2,3,4,5,6,7,8,9,10,11),4)) # create a vector of length 44
rep10 <- rep(deck[37:40],2) # create a bunch of 10s
deck <- sort(rep(c(deck,rep10),8)) # combine into one deck of 416 cards

twoplayerblackjack <- function(x){
  deald <- sample(x, 1, replace = F) # deal 1 card to dealer
  newdeck <- x[!x %in% deald] # remove those two cards from the deck so they can't be pulled again
  dd1 <- if (deald != 10){  # the above line removes all copies of the element, so we have to add the appropriate amount back in
    rep(deald,31) # if it isn't a 10, add it back 31 times (all 32 were removed, minus the one we pulled)
    } else{
    rep(deald, 95) # if it is a 10, add it back 95 times (same logic as above)
    }
  newdeck <- sort(c(newdeck, dd1))
  dealp <- sample(newdeck, 1, replace = F) # deal 1 card to player
    newdeck <- newdeck[!newdeck %in% dealp] # remove those two cards from the deck so they can't be pulled again
  pd1 <- if (dealp != 10){  # the above line removes all copies of the element, so we have to add the appropriate amount back in
    rep(dealp,31) # if it isn't a 10, add it back 31 times (all 32 were removed, minus the one we pulled)
    } else{
    rep(dealp, 95) # if it is a 10, add it back 95 times (same logic as above)
    }
  newdeck <- c(newdeck, pd1)
  deald <- sort(c(deald, sample(newdeck, 1, replace = F))) # create a vector of the first and second deal to dealer
   newdeck <- newdeck[!newdeck %in% deald[2]] # remove those two cards from the deck so they can't be pulled again
  dd2 <- if (deald[2] != 10){  # the above line removes all copies of the element, so we have to add the appropriate amount back in
    rep(deald[2],31) # if it isn't a 10, add it back 31 times (all 32 were removed, minus the one we pulled)
    } else{
    rep(deald[2], 95) # if it is a 10, add it back 95 times (same logic as above)
  } 
  newdeck <- sort(c(newdeck, dd2)) 
   dealp <- sort(c(dealp, sample(newdeck, 1, replace = F))) # create a vector of the first and second deal to player
    newdeck <- newdeck[!newdeck %in% dealp[2]] # remove those two cards from the deck so they can't be pulled again
  pd2 <- if (dealp[2] != 10){  # the above line removes all copies of the element, so we have to add the appropriate amount back in
    rep(dealp[2],31) # if it isn't a 10, add it back 31 times (all 32 were removed, minus the one we pulled)
    } else{
    rep(dealp[2], 95) # if it is a 10, add it back 95 times (same logic as above)
    }
  newdeck <- c(newdeck, pd2) # we now have two hands, but the deck won't update :(
  print(newdeck)
  add_dealer = deald[1] + deald[2]
  add_player = dealp[1] + dealp[2]
  print(add_dealer)
  print(add_player)
}

```


